Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report

### 2024-09-27 paper/paper.qmd
Minor change.
Code is styled using styler.

---
title: "Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report"
# subtitle:
author: 
  - Yizhe Chen
thanks: "Code and data are available at: https://github.com/YizheChenUT/Analyzing-Crime-Trends-in-Toronto-Insights-from-the-Police-Annual-Statistical-Report.git"
date: today
date-format: long
abstract: "This paper analyzes crime trends in Toronto using the Police Annual Statistical Report from 2014 to 2023. Data from Open Data Toronto is used to evaluate crime distributions across different categories and regions. Analysis reveals that some areas, particularly in divisions such as D13 and D32, report the highest number of property-related crimes like auto theft. Suburban areas also illustrate rising trends in specific crime subtypes. These findings can help optimize policing strategies and public safety resource allocations."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

## Plot crime trends by division
crime_by_division %>% 
  ggplot(aes(x = year, y = total_crimes, color = division_name)) +
  geom_line(size = 1) +
  labs(
    title = "Crime Incidents by Division (2014-2023)",
    x = "Year",
    y = "Total Incidents",
    color = "Division"
  ) +
  theme_minimal()# Minimal theme for clean visuals
```

@@

## Create a line chart showing crimes reported by year and crime category
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
crime_trends_plot %>%
  ggplot(aes(x = year, y = total_incidents, color = crime_category, group = crime_category)) +
  geom_line(size = 1) +  # Line plot for crime trends
  geom_point(size = 2) +  # Adding points to lines
  labs(
    title = "Crimes Reported by Year in Toronto (2014-2023)",
    x = "Year",
    y = "Total Incidents Reported",
    color = "Crime Category"
  ) +
  theme_minimal()   # Minimal theme for clean visuals

```

@@

## Create a bar chart comparing crime types and incidents by division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
crime_by_division_plot %>% 
  ggplot(aes(y = reorder(division_name, -total_incidents), x = total_incidents, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge") +  # Using a bar chart for comparison
  labs(
    title = "Crime Incidents by Division and Crime Type in Toronto",
    y = "Division",
    x = "Total Incidents",
    fill = "Crime Type"
  ) +
  theme_minimal() + # Minimal theme for clean visuals
  theme(
    axis.text.y = element_text(size = 8, margin = margin(r = 10)), # Add the spacing of y-axis
    plot.title = element_text(hjust = 0.5), # Placing title on the middle
    legend.position = "bottom"  # Placing legend on the bottom
  ) +
  guides(fill = guide_legend(ncol = 2))  # Making the legend 2 columns

```
Based on @fig-Crimes_by_Div, 32 Division reports high rates of auto theft than other divisions. 12 Division and 53 Division show large number reports in break-ins and residential thefts, especially in apartment.

@@

## Create a grouped bar chart to show clearance rate by crime type and division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
clearance_rate_plot %>%
  ggplot(aes(y = reorder(division_name, -avg_clearance_rate), x = avg_clearance_rate, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Grouped bar chart for clearance rates
  labs(
    title = "Clearance Rate by Crime Type and Division in Toronto",
    y = "Division",
    x = "Average Clearance Rate",
    fill = "Crime Type"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +  # Convert y-axis to percentage
  theme_minimal() +  # Minimal theme for clean visuals
  theme(
    axis.text.y = element_text(size = 12, margin = margin(r = 10)), # Add the spacing of y-axis
    plot.title = element_text(hjust = 0.5), # Placing title on the middle
    legend.position = "bottom"  # Placing legend on the bottom
  ) +
  guides(fill = guide_legend(ncol = 2))  # Making the legend 2 columns

```
As illustrated by @fig-Clearance_Rate, the overall clearance rate for property crimes remains low, with less than 10% of auto theft cases being resolved in most divisions. Crimes against persons, such as assault and sexual violations, generally have relatively high clearance rates, particularly in suburban areas.

>>>>

---
title: "Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report"
# subtitle:
author: 
  - Yizhe Chen
thanks: "Code and data are available at: https://github.com/YizheChenUT/Analyzing-Crime-Trends-in-Toronto-Insights-from-the-Police-Annual-Statistical-Report.git"
date: today
date-format: long
abstract: "This paper analyzes crime trends in Toronto using the Police Annual Statistical Report from 2014 to 2023. Data from Open Data Toronto is used to evaluate crime distributions across different categories and regions. Analysis reveals that some areas report the highest number of property-related crimes like auto theft. Suburban areas also illustrate rising trends in specific crime subtypes. These findings can help optimize policing strategies and public safety resource allocations."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

## Plot crime trends by division
crime_by_division %>%
  ggplot(aes(x = year, y = total_crimes, color = division_name)) +
  geom_line(size = 1) +
  labs(
    title = "Crime Incidents by Division (2014-2023)",
    x = "Year",
    y = "Total Incidents",
    color = "Division"
  ) +
  theme_minimal() # Minimal theme for clean visuals
```

@@

## Create a line chart showing crimes reported by year and crime category
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
crime_trends_plot %>%
  ggplot(aes(x = year, y = total_incidents, color = crime_category, group = crime_category)) +
  geom_line(size = 1) + # Line plot for crime trends
  geom_point(size = 2) + # Adding points to lines
  labs(
    title = "Crimes Reported by Year in Toronto (2014-2023)",
    x = "Year",
    y = "Total Incidents Reported",
    color = "Crime Category"
  ) +
  theme_minimal() # Minimal theme for clean visuals
```

@@

## Create a bar chart comparing crime types and incidents by division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
crime_by_division_plot %>%
  ggplot(aes(y = reorder(division_name, -total_incidents), x = total_incidents, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge") + # Using a bar chart for comparison
  labs(
    title = "Crime Incidents by Division and Crime Type in Toronto",
    y = "Division",
    x = "Total Incidents",
    fill = "Crime Type"
  ) +
  theme_minimal() + # Minimal theme for clean visuals
  theme(
    axis.text.y = element_text(size = 8, margin = margin(r = 10)), # Add the spacing of y-axis
    plot.title = element_text(hjust = 0.5), # Placing title on the middle
    legend.position = "bottom" # Placing legend on the bottom
  ) +
  guides(fill = guide_legend(ncol = 2)) # Making the legend 2 columns
```
Based on @fig-Crimes_by_Div, 32 Division reports high rates of auto theft than other divisions. 12 Division and 53 Division show large number reports in break-ins and residential thefts, especially in apartment.

@@

## Create a grouped bar chart to show clearance rate by crime type and division
## Citation: https://allisonhorst.github.io/palmerpenguins/
## Citation: https://ggplot2.tidyverse.org
clearance_rate_plot %>%
  ggplot(aes(y = reorder(division_name, -avg_clearance_rate), x = avg_clearance_rate, fill = crime_subtype)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) + # Grouped bar chart for clearance rates
  labs(
    title = "Clearance Rate by Crime Type and Division in Toronto",
    y = "Division",
    x = "Average Clearance Rate",
    fill = "Crime Type"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) + # Convert y-axis to percentage
  theme_minimal() + # Minimal theme for clean visuals
  theme(
    axis.text.y = element_text(size = 12, margin = margin(r = 10)), # Add the spacing of y-axis
    plot.title = element_text(hjust = 0.5), # Placing title on the middle
    legend.position = "bottom" # Placing legend on the bottom
  ) +
  guides(fill = guide_legend(ncol = 2)) # Making the legend 2 columns
```
As illustrated by @fig-Clearance_Rate, the overall clearance rate for property crimes remains low, with less than 10% of auto theft cases being resolved in most divisions. Crimes against persons, such as assault and sexual violations, generally have relatively high clearance rates, particularly in suburban areas.
